Non-word repetition in Yélî Dnye

### 2020-09-16 manuscript.Rmd
replied to mc comments

- add effect sizes to all analyses
- reply to questions

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether. 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->

@@

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.
<!--
A2M: do you agree with this decision or shall we do the alternative?
M2A: yeah, i think adding abirtrarily small frequencies to the NA cases is fraught for the reasons we talked about this morning (i.e., possible systematic undersampling in phoible that can't be helped).
-->

@@

subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.rounded is missing, use reported age
inc$age.rounded[is.na(inc$age.rounded)]<-inc$Age[is.na(inc$age.rounded)]
#summary(inc)
table(inc$birthOrder)->birthorder.tab
```

@@

#quick fix, but should be fixed upstream @ac
# M2A: Seems like you already did this, and the line below
# is setting the is.na(trnsc$Sex) cases to "G" -- I assume
# that's not a desirable outcome so am commenting it out
# a2m fixed, but I think it is better to fix all of these error upstream, in the demo.xlsx file
# M2A: Since it's fixed upstream I propose deleting it (and have therefore
# commented it out again for now)
# trnsc$Sex[trnsc$Sex=="F" & !is.na(trnsc$Sex)]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

@@

```{r overall}
# M2A: This excludes a bunch of kids (anyone with NA on any of these participant features),
# whittling the summary down to 24 from 42 -- is that intended?
# Also all phon_score is 100 and all NLD is 0; does that stike you as fishy?
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. Nearly all of them were: `r round(sum(!is.na(trnsc$meaning))/sum(!is.na(trnsc$mispronunciation))*100)`%.

**TO DISCUSS TOGETHER this is higher than any previous report! can we be sure of it?**

@@

```{r bylength}
# M2A: this reduces the total number of kids represented from 42 to 40
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children  (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p = `r corage$p.value`). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (`r tsex$statistic`) = `r tsex$parameter`, p = `r tsex$p.value`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, `r corbo$statistic`) = `r corbo$estimate`, p = `r corbo$p.value`) or maternal education  (data missing for `r sum(is.na(demo$matEd[demo$included=="yes"]))` children, `r corme$statistic`) = `r corme$estimate`, p = `r corme$p.value`)

>>>>

- add information about proportion of mispronunciations that are real based on other papers
- add effect sizes to all analyses
- ? correct treatment of words whose mispronunciation is an English word:
    mispronunciation meaning
361              two     two
362              two     two
879          no-pinêt                                          English: no peanut
908            nodedi                                           English: no daddy
933          no--dadi                                           English: no daddy
949           petrumi                                             English: "true"
950           petrumi                                             English: "true"
951           petrumi                                             English: "true"
1373              wee     English: where
1374              wee     English: where
1375              wee     English: where
1379              wee     English: where
1380              wee     English: where
1388              wee     English: where
1394              wee     English: where
1396              wee     English: where
1398              wee     English: where
1399              wee     English: where
1402              wat      English: what
1405              wat      English: what
1411              wee     English: where

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether (see Participants section for exclusions). 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->
<!-- a2m done -->

@@

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.

@@

#fix sex
demo$Sex[demo$Sex=="F"]<-"G"

subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.rounded is missing, use reported age
inc$age.rounded[is.na(inc$age.rounded)]<-inc$Age[is.na(inc$age.rounded)]
#summary(inc)
table(inc$birthOrder)->birthorder.tab

@@

trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

@@

```{r overall}
# M2A: This excludes a bunch of kids (anyone with NA on any of these participant features),
# whittling the summary down to 24 from 42 -- is that intended?
# Also all phon_score is 100 and all NLD is 0; does that stike you as fishy?
#a2m fixed and fixed!
aggregate(trnsc$correct*100,by=list(trnsc$id),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
overall=merge(overall, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(overall)<-c("ID","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. Two thirds of them were: `r round(sum(!is.na(trnsc$meaning) & trnsc$meaning != "0")/sum(!is.na(trnsc$mispronunciation))*100)`%.

@@

```{r bylength}
# M2A: this reduces the total number of kids represented from 42 to 40
#a2m fixed
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
by_length=merge(by_length, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(by_length)<-c("ID","Length","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children  (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p = `r corage$p.value`). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (`r tsex$statistic`) = `r tsex$parameter`, p = `r tsex$p.value`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, rho = (`r corbo$statistic`) = `r corbo$estimate`, p = `r corbo$p.value`) or maternal education  (data missing for `r sum(is.na(demo$matEd[demo$included=="yes"]))` children,  rho = ( `r corme$statistic`) = `r corme$estimate`, p = `r corme$p.value`).
